---
title: "Pilot2_Analysis_Report"
author: "Stephanie Bugler"
date: "2025-07-16"
output: html_document
---

```{r setup, include=FALSE}
#Set working directory for the entire file
knitr::opts_knit$set(root.dir = "/Users/stephaniebugler/Documents/Masters/Masters Thesis/R/Pilot_2/Data")

```

# Install Packages

```{r setup_packages, message=FALSE, warning=FALSE}
if (!require("pacman")) install.packages("pacman", repos = "https://cloud.r-project.org")
pacman::p_load(
  tidyverse,
  jsonlite,
  ggpubr,
  psych
)
```

# Load Libraries

```{r}
library(tidyverse)  
library(jsonlite)
library(ggpubr)
library(pacman)
library(psych)
```

# Load & Prep Data

```{r}

# Read the file line-by-line and parse each JSON row into a data frame
dat <- suppressMessages(
  read_file("Pilot2_data.txt") %>%
    str_split("\n") %>%
    first() %>%
    discard(~ .x == "" || .x == "\r") %>%
    map_dfr(fromJSON, flatten = TRUE)
)

# Assign unique ID to each participant
dat$ID <- NA
tmp_IDcounter <- 0
for (i in 1:nrow(dat)) {
  if (!is.na(dat$sender[i]) && dat$sender[i] == "Greetings") {
    tmp_IDcounter <- tmp_IDcounter + 1
  }
  dat$ID[i] <- tmp_IDcounter
}
rm(tmp_IDcounter)

```

# Score Calculations

```{r}

### ─────────────────────────────────────────────
### Calculate mean scores for Ecological Dimension
### ─────────────────────────────────────────────
# 1. Filter and select relevant columns
eco <- dat %>%
  filter(sender == "Ecological Dimension Scale") %>%
  select(ID, framingCondition, starts_with("EcologicalDimension"))

# 2. Convert char to numeric
eco[ , 3:ncol(eco)] <- lapply(eco[ , 3:ncol(eco)], function(x) as.numeric(trimws(x)))

# 3. Compute participant-level mean score
eco_scores <- eco
eco_scores$eco_mean <- rowMeans(eco_scores[, grep("EcologicalDimension", names(eco_scores))], na.rm = TRUE)
eco_scores <- eco_scores[, c("ID", "framingCondition", "eco_mean")]
```

```{r}
### ─────────────────────────────────────────────
### Bioinspiration scale: Reverse Coding + Mean Scores
### ─────────────────────────────────────────────

# 1. Extract raw Bioinspiration item responses
bio_items_raw <- dat %>%
  filter(sender == "Bioinspiration Scale") %>%
  select(ID, starts_with("Bioinspiration"))

# 2. Convert ALL columns except ID to numeric
bio_items_raw[ , 2:ncol(bio_items_raw)] <- lapply(bio_items_raw[ , 2:ncol(bio_items_raw)], function(x) as.numeric(trimws(x)))

# 3. Compute participant-level mean BEFORE reverse coding
bio_mean_before <- rowMeans(bio_items_raw[, -1], na.rm = TRUE)
overall_mean_before <- mean(bio_mean_before, na.rm = TRUE)

# 4. Reverse-code the 3 negatively worded items — FIXED to preserve ID
bio_items_clean <- dat %>%
  filter(sender == "Bioinspiration Scale") %>%
  select(ID, starts_with("Bioinspiration")) %>%
  mutate(
    across(starts_with("Bioinspiration"), ~ as.numeric(trimws(as.character(.)))),
    `Bioinspiration-IPI2r`  = 6 - `Bioinspiration-IPI2r`,
    `Bioinspiration-PN2r`   = 6 - `Bioinspiration-PN2r`,
    `Bioinspiration-VRtN4r` = 6 - `Bioinspiration-VRtN4r`
  )


# 5. Compute participant-level mean AFTER reverse coding
bio_mean_after <- rowMeans(bio_items_clean[, -1], na.rm = TRUE)
overall_mean_after <- mean(bio_mean_after, na.rm = TRUE)

# 5b. Assemble cleaned scores table
bio_scores <- bio_items_clean %>%
  mutate(bio_mean = bio_mean_after) %>%
  select(ID, bio_mean)


# 6. Print summary
cat("Average mean BEFORE reverse coding:", round(overall_mean_before, 3), "\n")
cat("Average mean AFTER reverse coding:", round(overall_mean_after, 3), "\n")


```

# Create table to compare framing conditions

```{r}

### ─────────────────────────────────────────────
### created framing info table
### ─────────────────────────────────────────────
# Extract from 'InformedConsent' page
frame_info <- dat %>%
  filter(sender == "InformedConsent") %>%
  select(ID, framingCondition)

# Merge true condition into score tables
eco_scores <- left_join(eco_scores, frame_info, by = "ID") %>%
  select(-framingCondition.x) %>%
  dplyr::rename(framingCondition = framingCondition.y)

bio_scores <- left_join(bio_scores, frame_info, by = "ID")


```

# 

# Summary & Plots

```{r}

#number of pps in each framing condition
table(eco_scores$framingCondition, useNA = "always")
```

```{r}

library(plyr)

# Summary function
data_summary <- function(data, varname, groupnames){
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm = TRUE),
      se = sd(x[[col]], na.rm = TRUE) / sqrt(length(na.omit(x[[col]]))))
  }
  data_sum <- ddply(data, groupnames, .fun = summary_func, varname)
  data_sum <- plyr::rename(data_sum, c("mean" = varname))
  return(data_sum)
}

# Summaries of participant-level scores by framing condition
eco_summary <- data_summary(eco_scores, "eco_mean", "framingCondition")
bio_summary <- data_summary(bio_scores, "bio_mean", "framingCondition")

# View results
print(eco_summary)
print(bio_summary)
```

# Interpretation of summary stats:

-   Sustainable framing leads to the highest ecological scores on average.

-   Sustainable framing produces the lowest bioinspiration scores.

-   Bioinspired framing leads to the highest bioinspiration scores.

-   Bioinspired framing results in the lowest ecological scores.

```{r}
# Plots the mean scores on the Bioinspiration & Ecological Dimension by Framing Condition
ggplot(eco_scores, aes(x = framingCondition, y = eco_mean)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Ecological Dimension by Framing", y = "Mean Score") +
  theme_minimal()

ggplot(bio_scores, aes(x = framingCondition, y = bio_mean)) +
  geom_boxplot(fill = "skyblue") +
  labs(title = "Bioinspiration by Framing", y = "Mean Score") +
  theme_minimal()
```

## Some more plots...

```{r}

# Violin plot for Ecological Dimension
ggplot(eco_scores, aes(x = framingCondition, y = eco_mean)) +
  geom_violin(fill = "palegreen", alpha = 0.5, trim = FALSE) +
  geom_jitter(width = 0.15, alpha = 0.4, color = "darkgreen") +
  stat_summary(fun = mean, geom = "point", color = "black", size = 3) +
  labs(title = "Ecological Scores by Framing", y = "Mean Score", x = "Framing Condition") +
  theme_minimal()

# Violin plot for Bioinspiration
ggplot(bio_scores, aes(x = framingCondition, y = bio_mean)) +
  geom_violin(fill = "lightblue", alpha = 0.5, trim = FALSE) +
  geom_jitter(width = 0.15, alpha = 0.4, color = "darkblue") +
  stat_summary(fun = mean, geom = "point", color = "black", size = 3) +
  labs(title = "Bioinspiration Scores by Framing", y = "Mean Score", x = "Framing Condition") +
  theme_minimal()
```

# Test For Spillover Effects

### i.e., Did exposure to the bioinspired vignettes results in pps. rating the tech as more sustainable? and vice versa.

For the Ecological Items/Sustainability Scale:

```{r}
# 1. Join pps bio inspiration and sustainability scores into one data frame
combined_scores <- left_join(bio_scores, eco_scores, by = c("ID", "framingCondition"))

#2. Run ANOVAs for spillover effects
#2a. Does framing affect Sustainability Scores?
eco_spillover <- aov(eco_mean ~ framingCondition, data = combined_scores)
summary(eco_spillover)
# 2a.2. Post-hoc comparision
TukeyHSD(eco_spillover)
```

Visualise the results for the Ecological Items/Sustainability Scale:

```{r}
#Visualise Framing effect on ecological scores
ggplot(combined_scores, aes(x = framingCondition, y = eco_mean, fill = framingCondition)) +
  geom_boxplot() +
  labs(title = "Spillover: Framing Effect on Ecological Scores",
       y = "Ecological Mean Score", x = "Framing Condition") +
  theme_minimal()
```

------------------------------------------------------------------------

For the Bioinspiration Scale:

```{r}

# Does framing affect bioinspiration scores?”
# post-hoc Tukey
bio_spillover <- aov(bio_mean ~ framingCondition, data = combined_scores)
summary(bio_spillover)
TukeyHSD(bio_spillover)
```

Visualise the results for the Bioinspiration Scale:

```{r}

#Visualise Framing effect on bioinpspiration scores
ggplot(combined_scores, aes(x = framingCondition, y = bio_mean, fill = framingCondition)) +
  geom_boxplot() +
  labs(title = "Spillover: Framing Effect on Bioinspiration Scores",
       y = "Bioinspiration Mean Score", x = "Framing Condition") +
  theme_minimal()

```

# Interpretation:

-   Sustainable vignettes significantly boosted sustainability attitudes (suggests the vignettes are working), but interestingly, bioinspired framing scored lowest on this scale.

-   This shows strong domain specificity — and possibly a negative spillover??

-   Bioinspired vignettes significantly increased bioinspiration scale scores i.e., the vignettes are working.

-   But again, sustainable framing actually reduced bioinspiration, suggesting minimal or even negative spillover in the other direction.

------------------------------------------------------------------------

## Everything below this point is broken

# Performance of Individual Scale Items

## Bioinspiration Scale

```{r}
# remove this awful package
remove.packages("plyr")
```

```{r}

 ### ─────────────────────────────────────────────
### Individual Item Analysis: Bioinspiration (Reversed) & Ecological Scales
### ─────────────────────────────────────────────

# 1. Convert cleaned + reverse-coded Bioinspiration data to long format
# Make sure the columns exist and are not dropped by mistake
bio_long <- bio_items_clean %>%
  pivot_longer(
    cols = starts_with("Bioinspiration"),
    names_to = "item",
    values_to = "score"
  ) %>%
  left_join(frame_info, by = "ID") %>%
  select(ID, item, score, framingCondition)  # <-- This is "immediately after your join"


# Check that items are back
glimpse(bio_long)
```

```{r}
bio_long %>% dplyr::count(.data$item, .data$framingCondition)

```

```{r}
# 4. Summarise by individual Bioinspiration item & framing condition


# bio_item_summary <- bio_long %>%
  #group_by(item, framingCondition) %>%
  #summarise(
    #mean = mean(score, na.rm = TRUE),
    #sd   = sd(score, na.rm = TRUE),
    #n    = sum(!is.na(score)),
    #se   = sd / sqrt(n),
    #.groups = "drop"
  #)

#glimpse(bio_item_summary)

```

```{r}
#  Convert Ecological Dimension data (cleaned) to long format
eco_joined <- eco %>%
  pivot_longer(
    cols = starts_with("EcologicalDimension"),
    names_to = "item",
    values_to = "score"
  ) %>%
  left_join(frame_info, by = "ID")

eco_long <- eco_joined %>%
  select(ID, item, score, framingCondition = framingCondition.y)



# Summarise by individual Ecological item & framing condition
eco_item_summary <- eco_long %>%
  group_by(item, framingCondition) %>%
  summarise(
    mean = mean(score, na.rm = TRUE),
    sd   = sd(score, na.rm = TRUE),
    n    = sum(!is.na(score)),
    se   = sd / sqrt(n),
    .groups = "drop"
  )

```

### Plot the Bioinspiration Scores by Individual Item

```{r}
# 3a. Plot Bioinspiration Scale items by framing

# First, reorder the items in a separate step
#bio_item_summary <- bio_item_summary %>%
  #mutate(item = fct_reorder(item, mean))

# Then plot
#ggplot(bio_item_summary, aes(x = item, y = mean, fill = framingCondition)) +
  #geom_col(position = position_dodge(width = 0.8)) +
  #geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                #width = 0.2, position = position_dodge(width = 0.8)) +
  #coord_flip() +
  #labs(
    #title = "Bioinspiration Items by Framing Condition",
    #x = "Item",
    #y = "Mean Score (1–5)"
  #) +
  #theme_minimal() +
  #theme(legend.position = "top")


```

# Ecological Dimension Scale

```{r}

# Create clean frame_info first (1 row per participant)
frame_info <- dat %>%
  filter(sender == "InformedConsent") %>%
  select(ID, framingCondition) %>%
  distinct()

# Now join to eco_long
eco_long <- eco %>%
  pivot_longer(
    cols = starts_with("EcologicalDimension"),
    names_to = "item",
    values_to = "score"
  ) %>%
  left_join(frame_info, by = "ID")

# Remove any existing 'framingCondition.x' and rename 'framingCondition.y'
eco_long <- eco_long %>%
  select(-framingCondition.x) %>%
  dplyr::rename(framingCondition = framingCondition.y)


# summarise eco items
eco_item_summary <- eco_long %>%
  group_by(item, framingCondition) %>%
  dplyr::summarise(
    mean = mean(score, na.rm = TRUE),
    sd = sd(score, na.rm = TRUE),
    n = sum(!is.na(score)),
    se = sd / sqrt(n),
    .groups = "drop"
  )

```

## Plot the Ecological Scale Items

```{r}

#visualise eco items
ggplot(eco_item_summary, aes(x = reorder(item, -mean), y = mean, fill = framingCondition)) +
  geom_col(position = position_dodge(width = 0.8)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                width = 0.2, position = position_dodge(width = 0.8)) +
  coord_flip() +
  labs(
    title = "Ecological Scale Items by Framing Condition",
    x = "Item", y = "Mean Score (1–7)"
  ) +
  theme_minimal() +
  theme(legend.position = "top")
```

------------------------------------------------------------------------

# Scale Purification for Bioinspiration Items

We want to find out which items performed well and which we can throw out.

### Run Cronbach's Alpha

```{r}

### ─────────────────────────────────────────────
### Bioinspiration Scale – Test for Internal Consistency
### ─────────────────────────────────────────────

# Load necessary libraries
library(dplyr)
library(psych)


# 1. Extract and clean Bioinspiration item responses
#bio_items_clean <- dat %>%
  #filter(sender == "Bioinspiration Scale") %>%
  #select(starts_with("Bioinspiration")) %>%
  #mutate(across(everything(), ~ as.numeric(trimws(as.character(.)))))  # Ensure numeric

# 2. Optional: unload plyr if loaded, to avoid conflicts
if ("package:plyr" %in% search()) {
  detach("package:plyr", unload = TRUE, character.only = TRUE)
}

# 3. Run Cronbach's alpha
bio_alpha <- alpha(bio_items_clean, check.keys = TRUE)  # Enables automatic detection of miskeyed items

# 4. Print results
cat("Cronbach's Alpha (raw):", round(bio_alpha$total$raw_alpha, 3), "\n\n")
print(bio_alpha$item.stats)

```

## 
